A radiocarbon data based evaluation of Cultural Transmission in Bronze Age burial rites of Central, Northern and North-western Europe

### 2019-01-20 article/04_Discussion.Rmd
a section about data meaning

Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:distanceandcorrelationtable*}). Only for small periods of time this correlation can be assumed based on the Radon-B data. The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems behave in a opposing manner: negative correlation. If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000calBC) and a decline (after 2000) of cultural-spatial-distance correlation. One interpretation of this oberservation might be, that the Early Bronze Age was a time of rising cultural exchange and religious homogenization, while in later periods different domains secluded themselves more strongly. There are two main reasons why such interpretations are difficult so far:

1. Data amount, distributiona and quality:

2. Data meaning

@@

>>>>

Against this background, the observation that there is no permanent, significant correlation either between the variables or with the spatial distance is all the more astonishing (Table \ref{tab:distanceandcorrelationtable*}). Only for small periods of time this correlation can be assumed based on the Radon-B data. The *burial construction* distance shows a higher degree of correlation with spatial distance than the *burial type* distance, which even seems to behave in an opposing manner: negative correlation. If one assumes the reconstructed timelines of cultural similarity intensity to be correct, than one has to explain a rise (between 2200 and 2000calBC) and a decline (after 2000) of cultural-spatial-distance correlation. One interpretation of this oberservation might be, that the Early Bronze Age was a time of rising cultural exchange and religious homogenization, while in later periods different domains secluded themselves more strongly. There are two main reasons why such interpretations are difficult so far:

1. The amount, distribution and quality of the data as discussed in section \ref{data-quality-discussion} queries the credibility of the cultural distance calculation.

2. Data meaning \ref{data-meaning-discussion}

@@

## Data meaning {#data-meaning-discussion}

This paper is based on ^14^C dates extracted from the Radon-B data set which were prepared for the dating of grave findings. ^14^C data on graves are used to chronologically attach a feature to the absolute chronological context or to better describe the relative development within a finding complex. The questions underlying the creation of a ^14^C date are therefore generally not equivalent to those imposed in this paper. While here the presence of an idea is to be measured abstractly at a certain place at a certain time, graves are usually dated exactly the other way round to contextualize and better understand them as concrete object. By definition, field research is oriented towards small-scale questions that can be answered to their respective extent. It must put a lot of energy into doing justice to all findings that are inevitably destroyed by excavation. There is a fundamental break between that great, theoretical archaeology, which wants to develop comprehensive narratives and only pay quantitative attention to the individual case, and the individual research project or emergency excavation, which has to dedicate a measure of respect and sometimes a lot of working time to each grave. In this thesis, ^14^C data are converted into a proxy of calibrated density distributions within artificial working regions, which is completely detached from the individual context.

The superordinate, large perspective also tries to simplify and categorize the parameters associated with a single burial. Radon-B anticipated this step by providing the variables *burial type* and *burial construction* that can be reduced to three categories each. This is extremely useful on the one hand, but also questionable on the other. The respective contexts can be multifarious: similar rituals in different cultural contexts can have different meanings. For this work, it is assumed -- with reference to *Cultural Evolution Theory* -- that despite this restriction, actions such as burning the dead or burying in burial mounds can be passed on across cultural and ethnic group boundaries through imitation and teaching. To put it provocatively: cremation burial in the Hungarian Early Bronze Age is a manifestation of the same idea as cremation in the Northern Late Bronze Age. This is a working hypothesis, but may be questioned.

The meaning of the data in Radon-B is thus manipulated in at least three ways for the purpose of this work: By concentrating on the dated grave in contrast to the ^14^C date, this work detaches itself from the actual content of the database. By creating artificial time steps and investigation regions, the cultural-historical context of the grave is dissolved. By creating the relative development proxies, the individual grave loses its context. The present paper is therefore a clear secondary use of Radon-B. A better dataset should be grave-related, not date-related, and should contain much more meta-information to each grave.
